Draft Final Report: Measles risk assessment, modelling, and benefit--cost analyses for New Zealand

### 2015-03-20 draftfinalreport_v2.Rnw
Major revisions to text

\documentclass{article}

\usepackage{times}
\usepackage{hyperref}
\usepackage{tikz}
%\usepackage{bibunits}
\usepackage{natbib}
\usepackage{graphics}
\usepackage{amsmath}
\usepackage{indentfirst}
\usepackage[utf8]{inputenc}
\usepackage{graphicx}
\usepackage{eurosym}
\usepackage{todonotes}
\usepackage{subfig}
\usepackage{enumitem}
\usepackage{pdflscape}
\usepackage{booktabs}
\usepackage{array}
\usepackage{rotating}
\usepackage{threeparttable}
\usepackage{blindtext}
\usepackage{dcolumn}
\usepackage{tabularx}
\usepackage{amssymb}
\usepackage{amsmath}
%\usepackage{caption2}
\usepackage{caption}
\usepackage{fancyhdr}
\usepackage{placeins}
\usepackage{framed}
\usepackage{float}
\usepackage{authblk}
\newcommand{\vecA}{\mathbf{A}}
\newcommand{\vecB}{\mathbf{B}}
\newcommand{\vecC}{\mathbf{C}}
\newcommand{\vecD}{\mathbf{D}}
\newcommand{\vecE}{\mathbf{E}}
\newcommand{\vecF}{\mathbf{F}}
\newcommand{\vecH}{\mathbf{H}}
\newcommand{\vecI}{\mathbf{I}}
\newcommand{\vecJ}{\mathbf{J}}
\newcommand{\vecK}{\mathbf{K}}
\newcommand{\vecM}{\mathbf{M}}
\newcommand{\vecN}{\mathbf{N}}
\newcommand{\vecP}{\mathbf{P}}
\newcommand{\vecQ}{\mathbf{Q}}
\newcommand{\vecS}{\mathbf{S}}
\newcommand{\vecT}{\mathbf{T}}
\newcommand{\vecW}{\mathbf{W}}
\newcommand{\vecZ}{\mathbf{Z}}
\newcommand{\vecb}{\mathbf{b}}
\newcommand{\vecc}{\mathbf{c}}
\newcommand{\vece}{\mathbf{e}}
\newcommand{\vecf}{\mathbf{f}}
\newcommand{\vecg}{\mathbf{g}}
\newcommand{\vech}{\mathbf{h}}
\newcommand{\veci}{\mathbf{i}}
\newcommand{\vecr}{\mathbf{r}}
\newcommand{\vecu}{\mathbf{u}}
\newcommand{\vecv}{\mathbf{v}}
\newcommand{\vecw}{\mathbf{w}}
\newcommand{\vecx}{\mathbf{x}}
\newcommand{\vecy}{\mathbf{y}}
\newcommand{\vecz}{\mathbf{z}}
\newcommand{\Ach}{\mathcal{H}}
\newcommand{\Ell}{\mathcal{L}}
\newcommand{\En}{\mathcal{N}}
\newcommand{\Pe}{\mathcal{P}}
\newcommand{\Ro}{\mathcal{R}_0}
\newcommand{\Rc}{\mathcal{R}_c}
\newcommand{\Pv}{\mathcal{P}_v}
\newcommand{\Rr}{\mathcal{R}}
\newcommand{\Qu}{\mathcal{Q}}
\newcommand{\Te}{\mathcal{T}}
\newcommand{\Ve}{\mathcal{V}}
\newcommand{\Dub}{\mathcal{W}}
\newcommand{\Zed}{\mathcal{Z}}
\newcommand{\Exp}{\mathbb{E}}
\newcommand{\de}{\mbox{d}}
\newcommand{\wrt}{\,\mbox{d}}
\newcommand{\veczero}{\mathbf{0}}
\newcommand{\vecone}{\mathbf{1}}
\newcommand{\vecalpha}{\mbox{\boldmath$\alpha$}}
\newcommand{\vectheta}{\mbox{\boldmath$\theta$}}
\newcommand{\vecphi}{\mbox{\boldmath$\phi$}}
\newcommand{\vecPhi}{\mbox{\boldmath$\Phi$}}
\newcommand{\vecSigma}{\mbox{\boldmath$\Sigma$}}
\newcommand{\vecimath}{\mbox{\boldmath$\imath$}}
\newcommand{\cosech}{\mbox{cosech}\,}
\newcommand{\sign}{\mbox{sign}\,}
\newcommand{\trace}{\mbox{trace}\,}
\newcommand{\Inf}{\mbox{Inf}\,}
%\newcommand{\Ro}{R_0}
\DeclareMathOperator{\var}{var}
\DeclareMathOperator{\cov}{cov}

@@

\begin{figure}
     %\centering
     \includegraphics[width=0.9\textwidth]{incidence_1997_2014.pdf}
     \caption{Measles incidence in New Zealand from 1997 to 2014}
     \label{fig:incidence1997}
\end{figure}

@@

<<echo=FALSE,include=FALSE,results=hide>>=
library(adehabitat)
library(classInt)
library(ggplot2)   # plotting library
library(grid)
library(gridExtra)
library(ggmap)
library(Hmisc)
library(maps)       # Provides functions that let us plot the maps
library(mapdata)    # Contains the hi-resolution points that mark out the countries.
library(maptools)
library(mapproj)
library(plyr)
library(pscl)
library(RColorBrewer)
library(rworldmap)
library(reshape)
library(reshape2)
library(rgeos)
library(rgdal)
library(raster)
library(sp)
library(utils)
library(xtable)
#library(SDMTools)
#gpclibPermit()

# read data
data<-read.csv("DHayman_20140627.csv",header=T)
vac<-read.csv("DHayman_20140715_Vacc.csv",header=T)
# names(data)
data$RptYear<-as.factor(data$RptYear)
data$SurvWeek<-as.factor(data$SurvWeek)
data$NZDep01<-as.factor(data$NZDep01)
data$NZDep06<-as.factor(data$NZDep06)
data$NZDep13<-as.factor(data$NZDep13)
#data$AgeInYears<-as.factor(data$AgeInYears)
data$EthnicityPrioritised<-as.factor(data$EthnicityPrioritised)

time<-aggregate( cbind( DiseaseName ) ~ NZDep01 +NZDep06+NZDep13+AgeInYears+EthnicityPrioritised + SurvWeek + RptYear,# + Dose1Mths + Dose2Mths, 
                 data = data , FUN=sum)
time_yr<-subset(time, (RptYear %in% c("2007","2008","2009","2010","2011","2012","2013","2014")))

caseyr<-aggregate( DiseaseName ~ AgeInYears, 
                   data = time_yr , FUN=sum)

popimmune<-read.csv("PopnImmunityAll.csv",header=T)
popimmune$Age = factor(popimmune$Age,levels(popimmune$Age)[c(2,3,6,8,10:12,4,5,7,9,1)])
pop<-read.csv("popnsize.csv",header=T)
colnames(pop)<-0:100
pop<-t(pop)
impop<-c(popimmune$Immunity[1:6],rep(popimmune$Immunity[7],8),rep(popimmune$Immunity[8],5),rep(popimmune$Immunity[9],5),
         rep(popimmune$Immunity[10],9),rep(popimmune$Immunity[11],20),rep(popimmune$Immunity[12],48))
naive<-round(pop-(pop*impop))

testv<-subset(timev, (RptYear %in% c("2007","2008","2009","2010","2011","2012","2013","2014")))

testv$Dose1Mths<-revalue(testv$Dose1Mths, c("9999"=NA));
testv$Dose2Mths<-revalue(testv$Dose2Mths, c("9999"=NA));
testv$Dose1Mths<-as.numeric(testv$Dose1Mths)
testv$Dose2Mths<-as.numeric(testv$Dose2Mths)
testv$D2vac<-ifelse(testv$Dose1Mths > -1 & testv$Dose2Mths >= testv$Dose1Mths,testv$Dose2Mths,NA)
testv$D1vac<-ifelse(testv$Dose1Mths > -1 & is.na(testv$D2vac) == T,testv$Dose1Mths,NA)
testv$Unvac<-ifelse(is.na(testv$Dose1Mths) == T & is.na(testv$Dose2Mths) == T,testv$AgeInYears,NA)

### denominator data
# setwd("~/Massey_2014/measles/data")
data<-read.csv("DHayman_20140627.csv",header=T)
vac<-read.csv("DHayman_20140715_Vacc.csv",header=T)
names(data)
data$RptYear<-as.factor(data$RptYear)
data$SurvWeek<-as.factor(data$SurvWeek)
data$NZDep01<-as.factor(data$NZDep01)
data$NZDep06<-as.factor(data$NZDep06)
data$NZDep13<-as.factor(data$NZDep13)
#data$AgeInYears<-as.factor(data$AgeInYears)
data$EthnicityPrioritised<-as.factor(data$EthnicityPrioritised)

test<-subset(time, (RptYear %in% c("2007","2008","2009","2010","2011","2012","2013","2014")))
caseyr<-aggregate( DiseaseName ~ AgeInYears, 
                   data = test , FUN=sum)
test$AgeInYears<-findInterval(test$AgeInYears,c(3,6,18,25))
test$AgeInYears<-as.factor(test$AgeInYears)
tage<-revalue(test$AgeInYears, c("0"="0-2", "1"="3-5","2"="6-17","3"="18-24","4"="25+"));
test$AgeInYears<-tage
#summary(test$EthnicityPrioritised)
teth<-revalue(test$EthnicityPrioritised, c("European or Other"="European", "Middle Eastern/Latin American/African"="MLA",
                                     "Pacific Peoples"="Pacific","Response cannot be classified"="None","Unknown"="None"));

test$EthnicityPrioritised<-teth

test <- within(test, NZDep06[NZDep06 == "0"]<-NA)
test <- within(test, NZDep13[NZDep13 == "0"]<-NA)

testt <- within(test, NZDep<- ifelse (test$RptYear == "2007",test$NZDep06,
                              ifelse (test$RptYear == "2008",test$NZDep06,
                              ifelse (test$RptYear == "2009",test$NZDep06,
                              ifelse (test$RptYear == "2010",test$NZDep06,
                              ifelse (test$RptYear == "2011",test$NZDep06,
                              ifelse (test$RptYear == "2012",test$NZDep06,
                              ifelse (test$RptYear == "2013",test$NZDep13,NZDep13))))))))
testt$NZDep<-as.factor(testt$NZDep-1)

testtable<-aggregate( cbind( DiseaseName) #+ as.numeric(SurvWeek))
                      ~ NZDep+AgeInYears+EthnicityPrioritised + RptYear, 
                 data = testt , FUN=sum)

colnames(testtable)<-c("NZDep","Age","Ethnicity","Year","Cases")

tnzd<-revalue(testtable$NZDep, c("1"="1-5","2"="1-5","3"="1-5","4"="1-5",
                                 "5"="1-5","6"="6-10","7"="6-10","8"="6-10",
                                 "9"="6-10","10"="6-10"));

testtable$NZDep<-tnzd

tt<-aggregate( cbind(Cases) ~ NZDep+Age+Ethnicity + Year, 
                 data = testtable , FUN=sum)

ttyr<-aggregate( cbind(Cases) ~ NZDep+Age+Ethnicity, 
                 data = testtable , FUN=sum)
denom<-read.csv("NZDep2006Denominators.csv",header=T)
denom<-denom[,-c(9)]

denom<-denom[,1:18]
row.has.na <- apply(denom, 1, function(x){any(is.na(x))})

denomt <- denom[!row.has.na,]

denomt<-denomt[denomt$Sex_code==99,]
denomt$Eth_Level<-as.factor(denomt$Eth_Level)
denomt<-denomt[denomt$Eth_Level %in% c("24","21","66",'22','77'),]
dm<-cbind(denomt[,c(3,6,9:18)])

dm<-dm[dm$Age_Label %in% c("A: 0- 5 yrs", "A: 6-17 yrs","A:18-24 yrs","A:25-64 yrs","A:65+ yrs"),]
dage<-revalue(dm$Age_Label, c("A: 0- 5 yrs"="<6", "A: 6-17 yrs"="6-17","A:18-24 yrs"="18-24","A:25-64 yrs"="25+","A:65+ yrs"="25+"));
dm$Age_Label<-dage
colnames(dm)<-c('Age','Ethnicity',"Dep1","Dep2","Dep3","Dep4",'Dep5','Dep6','Dep7','Dep8','Dep9',"Dep10")

datadm<-dm[dm$Age=="<6",]
dataA<-(rbind(datadm,datadm)/2)
dataA<-round(dataA[,3:12])
dataA<-cbind(c(rep("0-2",5),rep("3-5",5)),rep(datadm$Ethnicity[1:5],2),dataA)
colnames(dataA)<-c('Age','Ethnicity',"Dep1","Dep2","Dep3","Dep4",'Dep5','Dep6','Dep7','Dep8','Dep9',"Dep10")
ddm<-dm[!dm$Age=="<6",]
dm<-rbind(dataA,ddm)
dm$Ethnicity <- factor(dm$Ethnicity)
deth<-revalue(dm$Ethnicity, c("Asian (Prioritised)"="Asian","European (NZ European and Other European)"="European",
                              "Maori (Prioritised)"="Maori","MELAA"="MLA","Pacific People (Prioritised)"="Pacific"));
dm$Ethnicity<-deth
head(dm)

popn<-melt(dm,id.vars=c("Age","Ethnicity"),measure.vars=c("Dep1","Dep2","Dep3","Dep4",'Dep5','Dep6','Dep7','Dep8','Dep9',"Dep10"))
colnames(popn)<-c("Age","Ethnicity","NZDep","Popn")
pnzd<-revalue(popn$NZDep, c("Dep1"="1-5","Dep2"="1-5","Dep3"="1-5","Dep4"="1-5",
                                 "Dep5"="1-5","Dep6"="6-10","Dep7"="6-10","Dep8"="6-10",
                                 "Dep9"="6-10","Dep10"="6-10"));

popn$NZDep<-pnzd
tp<-aggregate( cbind(Popn) ~ NZDep+Age+Ethnicity, 
               data = popn , FUN=sum)

## get data
# set wd
# setwd("~/Massey 2014/DHayman_20140627")
# read data
data<-read.csv("DHayman_20140627.csv",header=T)
vac<-read.csv("DHayman_20140715_Vacc.csv",header=T)
names(data)
data$RptYear<-as.factor(data$RptYear)
data$SurvWeek<-as.factor(data$SurvWeek)
data$NZDep01<-as.factor(data$NZDep01)
data$NZDep06<-as.factor(data$NZDep06)
data$NZDep13<-as.factor(data$NZDep13)
#data$AgeInYears<-as.factor(data$AgeInYears)
data$EthnicityPrioritised<-as.factor(data$EthnicityPrioritised)

@@

testv<-subset(timev, (RptYear %in% c("2007","2008","2009","2010","2011","2012","2013","2014")))

testv$Dose1Mths<-revalue(testv$Dose1Mths, c("9999"=NA));
testv$Dose2Mths<-revalue(testv$Dose2Mths, c("9999"=NA));
testv$Dose1Mths<-as.numeric(testv$Dose1Mths)
testv$Dose2Mths<-as.numeric(testv$Dose2Mths)
testv$D2vac<-ifelse(testv$Dose1Mths > -1 & testv$Dose2Mths >= testv$Dose1Mths,testv$Dose2Mths,NA)
testv$D1vac<-ifelse(testv$Dose1Mths > -1 & is.na(testv$D2vac) == T,testv$Dose1Mths,NA)
testv$Unvac<-ifelse(is.na(testv$Dose1Mths) == T & is.na(testv$Dose2Mths) == T,testv$AgeInYears,NA)

@@

## regression analyses - measles
# setwd("~/Massey 2014/DHayman_20140627")
# read data
data<-read.csv("DHayman_20140627.csv",header=T)
vac<-read.csv("DHayman_20140715_Vacc.csv",header=T)
names(data)
data$RptYear<-as.factor(data$RptYear)
data$SurvWeek<-as.factor(data$SurvWeek)
data$NZDep01<-as.factor(data$NZDep01)
data$NZDep06<-as.factor(data$NZDep06)
data$NZDep13<-as.factor(data$NZDep13)
#data$AgeInYears<-as.factor(data$AgeInYears)
data$EthnicityPrioritised<-as.factor(data$EthnicityPrioritised)

@@

# test$AgeInYears<-findInterval(test$AgeInYears,c(3,6,18,25))
test$AgeInYears<-findInterval(test$AgeInYears,c(2,5,18,25))
test$AgeInYears<-as.factor(test$AgeInYears)
#tage<-revalue(test$AgeInYears, c("0"="0-2", "1"="3-5","2"="6-17","3"="18-24","4"="25+"));
tage<-revalue(test$AgeInYears, c("0"="0-1", "1"="2-4","2"="5-17","3"="18-24","4"="25+"));
test$AgeInYears<-tage
#summary(test$EthnicityPrioritised)
teth<-revalue(test$EthnicityPrioritised, c("European or Other"="European", "Middle Eastern/Latin American/African"="MLA",
"Pacific Peoples"="Pacific","Response cannot be classified"="None","Unknown"="None"));

test$EthnicityPrioritised<-teth

test <- within(test, NZDep06[NZDep06 == "0"]<-NA)
test <- within(test, NZDep13[NZDep13 == "0"]<-NA)

@@

@@

testtable$NZDep<-tnzd

tt<-aggregate( cbind(Cases) ~ NZDep+Age+Ethnicity + Year, 
data = testtable , FUN=sum)

ttyr<-aggregate( cbind(Cases) ~ NZDep+Age+Ethnicity, 
data = testtable , FUN=sum)
ttyr <- ttyr[ttyr$Ethnicity=="None",]
### denominator data

data<-read.csv("DHayman_20140627.csv",header=T)
vac<-read.csv("DHayman_20140715_Vacc.csv",header=T)

data$RptYear<-as.factor(data$RptYear)
data$SurvWeek<-as.factor(data$SurvWeek)
data$NZDep01<-as.factor(data$NZDep01)
data$NZDep06<-as.factor(data$NZDep06)
data$NZDep13<-as.factor(data$NZDep13)
#data$AgeInYears<-as.factor(data$AgeInYears)
data$EthnicityPrioritised<-as.factor(data$EthnicityPrioritised)

@@

caseyr<-aggregate( DiseaseName ~ AgeInYears, 
data = test , FUN=sum)
caseyr
## for MoH desired ages
test$AgeInYears<-findInterval(test$AgeInYears,c(2,5,18,25))
test$AgeInYears<-as.factor(test$AgeInYears)
tage<-revalue(test$AgeInYears, c("0"="0-1", "1"="2-4","2"="5-17","3"="18-24","4"="25+"));
test$AgeInYears<-tage
#summary(test$EthnicityPrioritised)
teth<-revalue(test$EthnicityPrioritised, c("European or Other"="European", "Middle Eastern/Latin American/African"="MLA",
"Pacific Peoples"="Pacific","Response cannot be classified"="None","Unknown"="None"));

@@

@@

@@

popn<-melt(dm,id.vars=c("Age","Ethnicity"),measure.vars=c("Dep1","Dep2","Dep3","Dep4",'Dep5','Dep6','Dep7','Dep8','Dep9',"Dep10"))
colnames(popn)<-c("Age","Ethnicity","NZDep","Popn")
popn

@@

tp$perCap<-with(tp,tp$cases/tp$Popn*10000)

tpsub<-tp[!(tp$Ethnicity=="MLA"),]
tpsub$ordert <- factor(tpsub$merge, as.character(tpsub$merge))

@@

tp$Ethnicity<- relevel(tp$Ethnicity, "European")
model<-glm(cases~Age*Ethnicity*NZDep+offset(log(Popn)),data=tp,family="quasipoisson")
summary(model)

@@

## use tpsub

@@

model3<-update(model2,~.-Ethnicity:NZDep)
summary(model3)
anova(model3,test="F")

@@

DOB<-as.numeric(levels(testv$RptYear))[testv$RptYear]-testv$AgeInYears
 hist(DOB,breaks=100,col="grey",main="2007- 2014 measles cases",xlab="Date of Birth")
 DOBVac<-table(testv$VC,DOB)
 row.names(DOBVac)<-c("Unvaccinated","Dose1","Dose2")
 DOBVac<-t(DOBVac)

popimmune<-read.csv("PopnImmunityAll.csv",header=T)
popimmune$Age = factor(popimmune$Age,levels(popimmune$Age)[c(2,3,6,8,10:12,4,5,7,9,1)])
pop<-read.csv("popnsize.csv",header=T)
colnames(pop)<-0:100
pop<-t(pop)
impop<-c(popimmune$Immunity[1:6],rep(popimmune$Immunity[7],8),rep(popimmune$Immunity[8],5),rep(popimmune$Immunity[9],5),
rep(popimmune$Immunity[10],9),rep(popimmune$Immunity[11],20),rep(popimmune$Immunity[12],48))
length(pop)
naive<-round(pop-(pop*impop))

popimmune<-read.csv("PopnImmunityAll.csv",header=T)
popimmune$Age = factor(popimmune$Age,levels(popimmune$Age)[c(2,3,6,8,10:12,4,5,7,9,1)])
pop<-read.csv("popnsize.csv",header=T)
colnames(pop)<-0:100
pop<-t(pop)
impop<-c(popimmune$Immunity[1:6],rep(popimmune$Immunity[7],8),rep(popimmune$Immunity[8],5),rep(popimmune$Immunity[9],5),
rep(popimmune$Immunity[10],9),rep(popimmune$Immunity[11],20),rep(popimmune$Immunity[12],48))
length(pop)
naive<-round(pop-(pop*impop))
## match cases per age
AgeInYears<-0:100
naive<-cbind(AgeInYears,naive)
colnames(naive)<-c("AgeInYears","Naive")
naive<-merge(naive,caseyr,by="AgeInYears",all=T)
#naive[is.na(naive)] <- 0
colnames(naive)<-c("AgeInYears","Naive","Cases")
##

>>>>

\documentclass{article}
\usepackage{longtable}
\usepackage{times}
\usepackage{hyperref}
\usepackage{tikz}
%\usepackage{bibunits}
\usepackage{natbib}
\usepackage{graphics}
\usepackage{amsmath}
\usepackage{indentfirst}
\usepackage[utf8]{inputenc}
\usepackage{graphicx}
\usepackage{eurosym}
\usepackage{todonotes}
\usepackage{subfig}
\usepackage{enumitem}
\usepackage{pdflscape}
\usepackage{booktabs}
\usepackage{array}
\usepackage{rotating}
\usepackage{threeparttable}
\usepackage{blindtext}
\usepackage{dcolumn}
\usepackage{tabularx}
\usepackage{amssymb}
\usepackage{amsmath}
%\usepackage{caption2}
\usepackage{caption}
\usepackage{fancyhdr}
\usepackage{placeins}
\usepackage{framed}
\usepackage{float}
\usepackage{authblk}
\newcommand{\vecA}{\mathbf{A}}
\newcommand{\vecB}{\mathbf{B}}
\newcommand{\vecC}{\mathbf{C}}
\newcommand{\vecD}{\mathbf{D}}
\newcommand{\vecE}{\mathbf{E}}
\newcommand{\vecF}{\mathbf{F}}
\newcommand{\vecH}{\mathbf{H}}
\newcommand{\vecI}{\mathbf{I}}
\newcommand{\vecJ}{\mathbf{J}}
\newcommand{\vecK}{\mathbf{K}}
\newcommand{\vecM}{\mathbf{M}}
\newcommand{\vecN}{\mathbf{N}}
\newcommand{\vecP}{\mathbf{P}}
\newcommand{\vecQ}{\mathbf{Q}}
\newcommand{\vecS}{\mathbf{S}}
\newcommand{\vecT}{\mathbf{T}}
\newcommand{\vecW}{\mathbf{W}}
\newcommand{\vecZ}{\mathbf{Z}}
\newcommand{\vecb}{\mathbf{b}}
\newcommand{\vecc}{\mathbf{c}}
\newcommand{\vece}{\mathbf{e}}
\newcommand{\vecf}{\mathbf{f}}
\newcommand{\vecg}{\mathbf{g}}
\newcommand{\vech}{\mathbf{h}}
\newcommand{\veci}{\mathbf{i}}
\newcommand{\vecr}{\mathbf{r}}
\newcommand{\vecu}{\mathbf{u}}
\newcommand{\vecv}{\mathbf{v}}
\newcommand{\vecw}{\mathbf{w}}
\newcommand{\vecx}{\mathbf{x}}
\newcommand{\vecy}{\mathbf{y}}
\newcommand{\vecz}{\mathbf{z}}
\newcommand{\Ach}{\mathcal{H}}
\newcommand{\Ell}{\mathcal{L}}
\newcommand{\En}{\mathcal{N}}
\newcommand{\Pe}{\mathcal{P}}
\newcommand{\Ro}{\mathcal{R}_0}
\newcommand{\Rc}{\mathcal{R}_c}
\newcommand{\Pv}{\mathcal{P}_v}
\newcommand{\Rr}{\mathcal{R}}
\newcommand{\Qu}{\mathcal{Q}}
\newcommand{\Te}{\mathcal{T}}
\newcommand{\Ve}{\mathcal{V}}
\newcommand{\Dub}{\mathcal{W}}
\newcommand{\Zed}{\mathcal{Z}}
\newcommand{\Exp}{\mathbb{E}}
\newcommand{\de}{\mbox{d}}
\newcommand{\wrt}{\,\mbox{d}}
\newcommand{\veczero}{\mathbf{0}}
\newcommand{\vecone}{\mathbf{1}}
\newcommand{\vecalpha}{\mbox{\boldmath$\alpha$}}
\newcommand{\vectheta}{\mbox{\boldmath$\theta$}}
\newcommand{\vecphi}{\mbox{\boldmath$\phi$}}
\newcommand{\vecPhi}{\mbox{\boldmath$\Phi$}}
\newcommand{\vecSigma}{\mbox{\boldmath$\Sigma$}}
\newcommand{\vecimath}{\mbox{\boldmath$\imath$}}
\newcommand{\cosech}{\mbox{cosech}\,}
\newcommand{\sign}{\mbox{sign}\,}
\newcommand{\trace}{\mbox{trace}\,}
\newcommand{\Inf}{\mbox{Inf}\,}
%\newcommand{\Ro}{R_0}
\DeclareMathOperator{\var}{var}
\DeclareMathOperator{\cov}{cov}

@@

\begin{sidewaysfigure}
     %\centering
     \includegraphics[width=0.9\textwidth]{incidence_1997_2014.pdf}
     \caption{Measles incidence in New Zealand from 1997 to 2014}
     \label{fig:incidence1997}
\end{sidewaysfigure}

@@

<<echo=FALSE,include=FALSE,results=hide>>=
rm(list=ls())
library(adehabitat)
library(classInt)
library(ggplot2)   # plotting library
library(grid)
library(gridExtra)
library(ggmap)
library(Hmisc)
library(maps)       # Provides functions that let us plot the maps
library(mapdata)    # Contains the hi-resolution points that mark out the countries.
library(maptools)
library(mapproj)
library(plyr)
library(pscl)
library(RColorBrewer)
library(rworldmap)
library(reshape)
library(reshape2)
library(rgeos)
library(rgdal)
library(raster)
library(sp)
library(utils)
library(xtable)
## get data
# set wd
# setwd("~/Massey 2014/DHayman_20140627")
# read data
data<-read.csv("DHayman_20140627.csv",header=T)
vac<-read.csv("DHayman_20140715_Vacc.csv",header=T)
names(data)
data$RptYear<-as.factor(data$RptYear)
data$SurvWeek<-as.factor(data$SurvWeek)
data$NZDep01<-as.factor(data$NZDep01)
data$NZDep06<-as.factor(data$NZDep06)
data$NZDep13<-as.factor(data$NZDep13)
#data$AgeInYears<-as.factor(data$AgeInYears)
data$EthnicityPrioritised<-as.factor(data$EthnicityPrioritised)

time<-aggregate( cbind( DiseaseName ) ~ NZDep01 +NZDep06+NZDep13+AgeInYears+EthnicityPrioritised + SurvWeek + RptYear,data = data , FUN=sum)

test<-subset(time, (RptYear %in% c("2007","2008","2009","2010","2011","2012","2013","2014")))
head(test)
dim(test)

@@

testv<-subset(timev, (RptYear %in% c("2007","2008","2009","2010","2011","2012","2013","2014")))
testv$Dose1Mths<-revalue(testv$Dose1Mths, c("9999"=NA));
testv$Dose2Mths<-revalue(testv$Dose2Mths, c("9999"=NA));
testv$Dose1Mths<-as.numeric(testv$Dose1Mths)
testv$Dose2Mths<-as.numeric(testv$Dose2Mths)
testv$D2vac<-ifelse(testv$Dose1Mths > -1 & testv$Dose2Mths >= testv$Dose1Mths,testv$Dose2Mths,NA)
testv$D1vac<-ifelse(testv$Dose1Mths > -1 & is.na(testv$D2vac) == T,testv$Dose1Mths,NA)
testv$Unvac<-ifelse(is.na(testv$Dose1Mths) == T & is.na(testv$Dose2Mths) == T,testv$AgeInYears,NA)

@@

## regression analyses - measles
# read data
data<-read.csv("DHayman_20140627.csv",header=T)
vac<-read.csv("DHayman_20140715_Vacc.csv",header=T)
names(data)
data$RptYear<-as.factor(data$RptYear)
data$SurvWeek<-as.factor(data$SurvWeek)
data$NZDep01<-as.factor(data$NZDep01)
data$NZDep06<-as.factor(data$NZDep06)
data$NZDep13<-as.factor(data$NZDep13)
#data$AgeInYears<-as.factor(data$AgeInYears)
data$EthnicityPrioritised<-as.factor(data$EthnicityPrioritised)

time<-aggregate( cbind( DiseaseName ) ~ NZDep01 +NZDep06+NZDep13+AgeInYears+EthnicityPrioritised + SurvWeek + RptYear, data = data , FUN=sum)

@@

test$AgeInYears<-findInterval(test$AgeInYears,c(2,5,18,25))
test$AgeInYears<-as.factor(test$AgeInYears)
tage<-revalue(test$AgeInYears, c("0"="0-1", "1"="2-4","2"="5-17","3"="18-24","4"="25+"));
test$AgeInYears<-tage
teth<-revalue(test$EthnicityPrioritised, c("European or Other"="European", "Middle Eastern/Latin American/African"="MLA","Pacific Peoples"="Pacific","Response cannot be classified"="None","Unknown"="None"));

test$EthnicityPrioritised<-teth
test <- within(test, NZDep06[NZDep06 == "0"]<-NA)
test <- within(test, NZDep13[NZDep13 == "0"]<-NA)

@@

testtable<-aggregate( cbind( DiseaseName) ~ NZDep+AgeInYears+EthnicityPrioritised + RptYear,  data = testt , FUN=sum)
dim(testtable)
colnames(testtable)<-c("NZDep","Age","Ethnicity","Year","Cases")
head(testtable)

@@

ttyr<-aggregate( cbind(Cases) ~ NZDep+Age+Ethnicity, 
data = testtable , FUN=sum)
ttyr <- ttyr[ttyr$Ethnicity=="None",]

# latex(ttyr, file="",table.env=FALSE,rowname=NULL)

### denominator data
data<-read.csv("DHayman_20140627.csv",header=T)
vac<-read.csv("DHayman_20140715_Vacc.csv",header=T)
names(data)
data$RptYear<-as.factor(data$RptYear)
data$SurvWeek<-as.factor(data$SurvWeek)
data$NZDep01<-as.factor(data$NZDep01)
data$NZDep06<-as.factor(data$NZDep06)
data$NZDep13<-as.factor(data$NZDep13)
#data$AgeInYears<-as.factor(data$AgeInYears)
data$EthnicityPrioritised<-as.factor(data$EthnicityPrioritised)

@@

test<-subset(time, (RptYear %in% c("2007","2008","2009","2010","2011","2012","2013","2014")))
head(test)
dim(test)
caseyr<-aggregate( DiseaseName ~ AgeInYears, 
data = test , FUN=sum)
caseyr
## for MoH desired ages
test$AgeInYears<-findInterval(test$AgeInYears,c(2,5,18,25))
test$AgeInYears<-as.factor(test$AgeInYears)
tage<-revalue(test$AgeInYears, c("0"="0-1", "1"="2-4","2"="5-17","3"="18-24","4"="25+"));
test$AgeInYears<-tage
#summary(test$EthnicityPrioritised)
teth<-revalue(test$EthnicityPrioritised, c("European or Other"="European", "Middle Eastern/Latin American/African"="MLA",
"Pacific Peoples"="Pacific","Response cannot be classified"="None","Unknown"="None"));

@@

testt <- within(test, NZDep<- ifelse (test$RptYear == "2007",test$NZDep06,
ifelse (test$RptYear == "2008",test$NZDep06,
ifelse (test$RptYear == "2009",test$NZDep06,
ifelse (test$RptYear == "2010",test$NZDep06,
ifelse (test$RptYear == "2011",test$NZDep06,
ifelse (test$RptYear == "2012",test$NZDep06,
ifelse (test$RptYear == "2013",test$NZDep13,NZDep13))))))))
head(testt)
tail(testt)
str(testt)
testt$NZDep<-as.factor(testt$NZDep-1)

@@

#tnzd<-revalue(testtable$NZDep, c("1"="1-5","2"="1-5","3"="1-5","4"="1-5",
#"5"="1-5","6"="6-10","7"="6-10","8"="6-10",
#"9"="6-10","10"="6-10"));

@@

popn<-melt(dm,id.vars=c("Age","Ethnicity"),measure.vars=c("Dep1","Dep2","Dep3","Dep4",'Dep5','Dep6','Dep7','Dep8','Dep9',"Dep10"))
colnames(popn)<-c("Age","Ethnicity","NZDep","Popn")
popn
#pnzd<-revalue(popn$NZDep, c("Dep1"="1-5","Dep2"="1-5","Dep3"="1-5","Dep4"="1-5",
#"Dep5"="1-5","Dep6"="6-10","Dep7"="6-10","Dep8"="6-10",
#"Dep9"="6-10","Dep10"="6-10"));

@@

tp$perCap<-with(tp,tp$cases/tp$Popn*10000)
tpsub<-tp[!(tp$Ethnicity=="MLA"),]
tpsub$ordert <- factor(tpsub$merge, as.character(tpsub$merge))

@@

#hist(tp$cases,xlab="Cases",main='Histogram of cases per category',col='grey',breaks=20)
tp$Ethnicity<- relevel(tp$Ethnicity, "European")
model<-glm(cases~Age*Ethnicity*NZDep+offset(log(Popn)),data=tp,family="quasipoisson")
summary(model)

@@

## use tpsub
#hist(tpsub$cases,xlab="Cases",main='Histogram of cases per category',col="darkgrey",breaks=20)

@@

model3<-update(model2,~.-Ethnicity:NZDep)
summary(model3)
anova(model3,test="F")
anovap<-anova(model3,test="F")

res<-predict(model3)
#res<-predict(model2)

cor(exp(res),tpsub$cases)
cor.test(exp(res),tpsub$cases)

resan<-xtable(anova(model3,test="F"))
#resan<-xtable(anova(model2,test="F"))
# print(resan,floating = FALSE)

#xtb<-xtable(summary(model3))
xtb<-xtable(summary(model2))
# print(xtb,floating = FALSE)

@@

DOB<-as.numeric(levels(testv$RptYear))[testv$RptYear]-testv$AgeInYears
#hist(DOB,breaks=100,col="grey",main="2007- 2014 measles cases",xlab="Date of Birth")

DOBVac<-table(testv$VC,DOB)
row.names(DOBVac)<-c("Unvaccinated","Dose1","Dose2")
DOBVac<-t(DOBVac)

popimmune<-read.csv("PopnImmunityAll.csv",header=T)
popimmune$Age = factor(popimmune$Age,levels(popimmune$Age)[c(2,3,6,8,10:12,4,5,7,9,1)])
pop<-read.csv("popnsize.csv",header=T)
colnames(pop)<-0:100
pop<-t(pop)
impop<-c(popimmune$Immunity[1:6],rep(popimmune$Immunity[7],8),rep(popimmune$Immunity[8],5),rep(popimmune$Immunity[9],5),
rep(popimmune$Immunity[10],9),rep(popimmune$Immunity[11],20),rep(popimmune$Immunity[12],48))
length(pop)
naive<-round(pop-(pop*impop))
## match cases per age
AgeInYears<-0:100
naive<-cbind(AgeInYears,naive)
colnames(naive)<-c("AgeInYears","Naive")
naive<-merge(naive,caseyr,by="AgeInYears",all=T)
#naive[is.na(naive)] <- 0
colnames(naive)<-c("AgeInYears","Naive","Cases")
##
###
